\bf Nonparametric Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test for Marginal Distributions of Stationary Time Series

### 2024-03-14 manuscript/manuscript.tex
little smoothing in introductions and methods

The Kolmogorov-Smirnov (KS) statistic is widely used to test if a sample is
from a given distribution. This study demonstrates how non-parametric 
block bootstrap can be 
used to approximate the KS statistic in situations where parameters are 
not specified and the data are serially dependent. We demonstrate the 
theoretical justication for this method, then through simulation we verify that
it holds it size under the null distribution and is powerful when the null 
hypothesis is false. Finally, we demonstrate applications of the method to
precipitation data from three airports and Microsoft stock returns.

@@

Despite its widespread use, the Kolmogorov-Smirnov (KS) test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines 
common misapplications of the one-sample KS test, notably the inappropriate
use when the hypothesized distribution contains unspecified parameters.
For this situation, \citet{zeimbekakis2022misuses} proposed a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from samples generated from the fitted hypothesized distribution.
Another commonly encountered situation is when the data are serially
dependent. \citet{babu2004goodness} addresses this issue through the
application of basic non-parametric bootstrap methods. Alternatively, 
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

Developing a fully non-parametric solution for the Kolmogorov-Smirnov (KS) 
tests in the presence of serially dependent data presents a significant 
challenge. The null distribution's characteristics are intricately linked to 
the structure of the serial dependence, which can vary widely in practical 
situations. Employing a parametric bootstrap would necessitate defining a 
specific model for this dependence, despite the primary focus being on 
evaluating the marginal distribution of a stationary series. To date, a bias 
correction method for block bootstrap, akin to the approach \citet{babu2004goodness} 
applied to the basic bootstrap, has not been established. This study seeks 
to bridge this gap by introducing a bias correction technique for the 
non-parametric block bootstrap, tailored for use in scenarios where the 
hypothesized distribution includes unspecified parameters and the data 
exhibit serial dependence. Our approach reduces to that of
\citet{babu2004goodness} when the data are independent.

The remainder of this paper is structured as follows: 
Section~\ref{sec:methods} provides an overview of the block bootstrap 
procedure and introduces the bias correction methodology for the 
non-parametric block bootstrap KS test. Section~\ref{sec:simu} is divided 
into two parts; initially, we evaluate the KS test's ability to maintain 
its size, i.e., its consistency in not rejecting the null hypothesis when 
it is indeed true. Subsequently, we examine the test's power, assessing 
whether it can reject the null hypothesis when it is false. Practical 
applications of our method are presented in Section~\ref{sec:real}, where 
we first apply the approach to assess whether annual maximums of hourly 
precipitation at three U.S. airports conform to the Generalized Extreme 
Value (GEV) distribution in Subsection~\ref{sec:precipitation}. Following 
this, we analyze if Microsoft stock return data adhere to either the Normal 
or the Student's $t$ distribution in Subsection~\ref{sec:microsoft}. The 
paper concludes with Section~\ref{sec:conclusion}, offering final thoughts 
and remarks.

@@

Consider a stationary time series $\{X_t: t = 1, \ldots, n\}$ with length~$n$.
We are interested in testing whether or not $X_t$ follows a distribution in a
parametric family of distribution~$F$ indexed by a parameter
vector~$\theta$. That is, the null hypothesis is
\[
  H_0: X_t \sim F(\cdot \mid \theta), \quad t = 1, \ldots, n,
\]
for some unspecified parameter $\theta$.
The alternative hypothesis $H_A$ is that the marginal distribution of $X_t$ does
not follow~$F$ for any parameter value~$\theta$. This is a challenging situation
both the parameters and the serial dependence structure are unknown.

@@

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1,...,X^{(b)}_n$ are generated by block bootstrap for 
$b \in \{1, \ldots, B\}$.  
Block-bootstrap can be done with overlapping or moving blocks.
Define 
moving blocks (assuming $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}
A common 
function for block size that is considered optimal is 
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block},  
which was adopted in this study.

We then use bootstrap to approximate the distribuition of
the KS statistic under $H_0$. In this case, we can 
use the distribution of $\sqrt{n}(F^{(b)}_n(x) - E[F^{(b)}_n(x)])$
as an approximation of the distribution of
$\sqrt{n}(F_n(x) - F(x))$,
and the distribution of 
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E[\hat\theta^{(b)}_n]))$ to
approximate $\sqrt{n}(F(x; \theta_n) - F(x))$'s distribution.

>>>>

The Kolmogorov-Smirnov (KS) statistic is widely used to test if a sample is
from a given distribution. This study demonstrates how non-parametric 
block bootstrap can be 
used to approximate the KS statistic in situations where parameters are 
not specified and the data are serially dependent. We demonstrate the 
theoretical justication for this method, then through simulation we verify that
it holds it size under the null distribution and is powerful when the null 
hypothesis is false. Finally, we demonstrate applications of the method to
the S&P 100 and S&P 500 indices stock returns.

@@

Despite its widespread use, the KS test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines 
common misapplications of the one-sample KS test, notably the inappropriate
use when the hypothesized distribution contains unspecified parameters.
For this situation, \citet{zeimbekakis2022misuses} proposed a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from samples generated from the fitted hypothesized distribution.
Alternatively, \citet{babu2004goodness} addresses this issue through the
application of basic non-parametric bootstrap methods.
Another commonly encountered situation is when the data are serially
dependent.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

Developing a fully non-parametric solution for the KS 
tests in the presence of serially dependent data presents a significant 
challenge. The null distribution's characteristics are intricately linked to 
the structure of the serial dependence, which can vary widely in practical 
situations. Employing a non-parametric bootstrap would necessitate defining a 
specific model for this dependence, despite the primary focus being on 
evaluating the marginal distribution of a stationary series. To date, a bias 
correction method for block bootstrap, akin to the approach \citet{babu2004goodness} 
applied to the basic bootstrap, has not been established. This study seeks 
to bridge this gap by introducing a bias correction technique for the 
non-parametric block bootstrap, tailored for use in scenarios where the 
hypothesized distribution includes unspecified parameters and the data 
exhibit serial dependence. Our approach reduces to that of
\citet{babu2004goodness} when the data are independent.

The remainder of this paper is structured as follows: 
Section~\ref{sec:methods} provides an overview of the block bootstrap 
procedure and introduces the bias correction methodology for the 
non-parametric block bootstrap KS test. Section~\ref{sec:simu} is divided 
into two parts; initially, we evaluate the KS test's ability to maintain 
its size, i.e., its consistency in not rejecting the null hypothesis when 
it is indeed true. Subsequently, we examine the test's power, assessing 
whether it can reject the null hypothesis when it is false. Practical 
applications of our method are presented in Section~\ref{sec:real}, where 
we apply the approach to analyze if S&P 100 and S&P 500 index stock return data adhere to either the Normal 
or the Student's $t$ distribution in Subsection~\ref{sec:sp100} and
\ref{sec:sp500}. The 
paper concludes with Section~\ref{sec:conclusion}, offering final thoughts 
and remarks.

@@

Consider a stationary time series $\{X_t: t = 1, \ldots, n\}$ with length~$n$.
We are interested in testing whether or not $X_t$ follows a distribution in a
parametric family of distribution~$F$ indexed by a parameter
vector~$\theta$. That is, the null hypothesis is
\[
  H_0: X_t \sim F(\cdot \mid \theta), \quad t = 1, \ldots, n,
\]
for some unspecified parameter $\theta$.
The alternative hypothesis $H_A$ is that the marginal distribution of $X_t$ does
not follow~$F$ for any parameter value~$\theta$. This is a challenging situation
because both the parameters and the serial dependence structure are unknown.

@@

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1,...,X^{(b)}_n$ are generated by block bootstrap for 
$b \in \{1, \ldots, B\}$.  
Block-bootstrap can be done with overlapping or moving blocks.
Define 
moving blocks (assuming $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}
A common 
function for block size that is considered optimal is 
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block},  
which was adopted in this study.
Now we draw $k$ blocks from the $(n - l + 1)$ blocks 
of $Z_j$'s with replacement and then align them in the order they were picked to
form a block bootstrap sample. If $n$ is not a multiple of~$l$, the last block 
selected will be reduced in size so that the final size of the block bootstrap 
sample is $n$.

We then use bootstrap to approximate the distribution of
the KS statistic under $H_0$. In this case, we can 
use the distribution of $\sqrt{n}(F^{(b)}_n(x) - E[F^{(b)}_n(x)])$
as an approximation of the distribution of
$\sqrt{n}(F_n(x) - F(x))$,
and the distribution of 
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E[\hat\theta^{(b)}_n]))$ to
approximate $\sqrt{n}(F(x; \theta_n) - F(x))$'s distribution.
